Imputation of Incomplete Multilevel Data with \pkg{mice}

### 2022-12-13 Manuscript/Imputation_of_Incomplete_Multilevel_Data.tex

Table 1: Concepts in multilevel methods \textbar{} \textbf{Concept}
\textbar{} \textbf{Details} \textbar{}
\textbar---------------------\textbar-------------------------------------------------------------------------------\textbar{}
\textbar{} Sample unit \textbar{} Units of the population from which
measurements are taken in a sample. \textbar{} \textbar{} Hierarchical
levels \textbar{} Data are grouped into clusters at different levels. A
three-level \textbar{} \textbar{} Fixed effect \textbar{} Here we assume
that the values of an independent variable are fixed, i.e., \textbar{}
\textbar{} \textbar{} the values observed in the study are
representative of all values in the \textbar{} \textbar{} \textbar{} in
the dependent variable y e.g., blood pressure between treatments A and
B. \textbar{} \textbar{} Random effect \textbar{} The values of an
independent variable are assumed to be randomly drawn from \textbar{}
\textbar{} \textbar{} admission we might select only certain hospitals
that are representative of \textbar{} \textbar{} \textbar{} the
difference of y between individual hospitals, but rather the variation
of \textbar{} \textbar{} ICC \textbar{} The variability due to
clustering is often measured by means of the \textbar{} \textbar{}
\textbar{} intraclass coefficient (ICC). The ICC can be seen as the
percentage \textbar{} \textbar{} \textbar{} of variance that can be
attributed to the cluster-level, where a high \textbar{} \textbar{}
\textbar{} ICC would indicate that a lot of variability is due to the
cluster \textbar{} \textbar{} \textbar{} structure. \textbar{}
\textbar{} Random effect \textbar{} Multilevel models typically
accommodate for variability by including \textbar{} \textbar{}
\textbar{} a separate group mean for each cluster. In addition to random
\textbar{} \textbar{} \textbar{} intercepts, multilevel models can also
include random coefficients \textbar{} \textbar{} \textbar{} and
heterogeneous residual error variances across clusters \citep[see
e.g.~\textbar{} \textbar{} \textbar{}][\citet{hox17} and
\citet{jong21}]{gelm06}. {[}TODO: add stratification.{]} \textbar{}

@@

Table 2: Concepts in missing data methods \textbar{} \textbf{Concept}
\textbar{} \textbf{Details} \textbar{}
\textbar----------------\textbar----------------------------------------------------------------------------\textbar{}
\textbar{} MCAR \textbar{} Missing Completely At Random, where the
probability to be missing is equal \textbar{} \textbar{} \textbar{}
across all data entries \textbar{} \textbar{} MAR \textbar{} Missing At
Random, where the probability to be missing depends on observed
\textbar{} \textbar{} \textbar{} information \textbar{} \textbar{} MNAR
\textbar{} Missing Not At Random (MNAR), where the probability to be
missing \textbar{} \textbar{} \textbar{} depends on unrecorded
information, making the missingness non-ignorable \textbar{} \textbar{}
\textbar{} \citep{rubi76, meng94}. \textbar{} \textbar{} \textbar{}
{[}TODO: add congeniality, but maybe in-text?{]} \textbar{}

>>>>

Table 1: Concepts in multilevel methods

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.2100}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.7900}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Concept}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Details}
\end{minipage} \\
\midrule()
\endhead
Sample unit & Units of the population from which measurements are taken
in a sample. \\
Hierarchical levels & Data are grouped into clusters at different
levels. A three-level \\
Fixed effect & Here we assume that the values of an independent variable
are fixed, i.e., \\
& the values observed in the study are representative of all values in
the \\
& in the dependent variable y e.g., blood pressure between treatments A
and B. \\
Random effect & The values of an independent variable are assumed to be
randomly drawn from \\
& admission we might select only certain hospitals that are
representative of \\
& the difference of y between individual hospitals, but rather the
variation of \\
ICC & The variability due to clustering is often measured by means of
the \\
& intraclass coefficient (ICC). The ICC can be seen as the percentage \\
& of variance that can be attributed to the cluster-level, where a
high \\
& ICC would indicate that a lot of variability is due to the cluster \\
& structure. \\
Random effect & Multilevel models typically accommodate for variability
by including \\
& a separate group mean for each cluster. In addition to random \\
& intercepts, multilevel models can also include random coefficients \\
& and heterogeneous residual error variances across clusters {[}see
e.g. \\
& \citet{gelm06}, \citet{hox17} and \citet{jong21}{]}. {[}TODO: add
stratification.{]} \\
\bottomrule()
\end{longtable}

@@

Table 2: Concepts in missing data methods

\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.1739}}
  >{\raggedright\arraybackslash}p{(\columnwidth - 2\tabcolsep) * \real{0.8261}}@{}}
\toprule()
\begin{minipage}[b]{\linewidth}\raggedright
\textbf{Concept}
\end{minipage} & \begin{minipage}[b]{\linewidth}\raggedright
\textbf{Details}
\end{minipage} \\
\midrule()
\endhead
MCAR & Missing Completely At Random, where the probability to be missing
is equal \\
& across all data entries \\
MAR & Missing At Random, where the probability to be missing depends on
observed \\
& information \\
MNAR & Missing Not At Random (MNAR), where the probability to be
missing \\
& depends on unrecorded information, making the missingness
non-ignorable \\
& \citep{rubi76, meng94}. \\
& {[}TODO: add congeniality, but maybe in-text?{]} \\
\bottomrule()
\end{longtable}
